---
title: "Vignette"
output: rmarkdown::html_vignette
author:
  - "Grace Campidilli (gec83)"
  - "Benjamin Yeh (by253)"
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# load all functions from within this package
library(atnPkg)

# load external packages
library(fields)
library(maps)
library(sp)
library(ggplot2)

# Get unique airports ID
airports = airport_data$codes
n_airports = length(airports)
```


1. Make a map of the airport data.
```{r fig.height=4, fig.width=6}
airport_dat_for_map = data.frame(lon = airport_data$lon, lat = airport_data$lat)
rownames(airport_dat_for_map) = airport_data$codes

# Generate all possible trips (m=2)- i.e. every possible starting point mapped to every possible ending point
all_trips = as.data.frame(cbind(t(combn(airport_data$lon, 2)), t(combn(airport_data$lat, 2))))
colnames(all_trips) = c("lon1","lon2","lat1","lat2")

# Create text labels for map that have airport code and number of direct flights
airport_info = unlist(lapply(1:length(direct_from), function(x){
  paste(airport_data$codes[x],' (',length(direct_from[[airports[x]]]),')', sep = '')
}))

# plot map of airport data
par(mar=c(0,0,0,0))
map('world',col="#f2f2f2",fill=TRUE, bg="white", lwd=0.05,mar=rep(0,4),border=0, ylim=c(-80,80) )
points(x=airport_data$lon, y=airport_data$lat, col="slateblue", cex=0.25, pch=20)
text(airport_info, x=airport_data$lon, y=airport_data$lat,  col="slateblue", cex=0.15, pos = 4, offset = 1/5)

```

Below is code that defines a grid of points covering the continental USA
and figures out which points are inside the continental USA.

```{r}
# Get the polygon defining the continental USA
# usa_poly <- map("usa")
# usa_poly$x <- c(NA,usa_poly$x,NA)
# usa_poly$y <- c(NA,usa_poly$y,NA)
# nai <- which( is.na( usa_poly$x ) )
# 
# # Define a grid of lonitude and latitude points
# n1 <- 180
# n2 <- 90
# lo <- seq(usa_poly$range[1], usa_poly$range[2], length.out = n1)
# la <- seq(usa_poly$range[3], usa_poly$range[4], length.out = n2)
# lonlat_grid <- as.matrix( expand.grid( lo, la ) )
# 
# # Figure out which points are inside USA
# in_usa <- rep(FALSE, nrow(lonlat_grid))
# for(j in 1:(length(nai)-1)){
#   in_this <- sp::point.in.polygon(
#     lonlat_grid[,1],
#     lonlat_grid[,2],
#     usa_poly$x[ (nai[j]+1):(nai[j+1]-1) ],
#     usa_poly$y[ (nai[j]+1):(nai[j+1]-1) ]
#   )
#   in_usa <- in_usa | in_this
# }
# 
# # Subset to the points in USA
# lonlat_usa <- as.matrix( lonlat_grid[ in_usa, ] )
# # Naming
# colnames(lonlat_usa) = c('lon', 'lat')
# lonlat_usa = data.frame(lonlat_usa)
n_points = nrow(lonlat_usa)
```


2. Given a point on the grid, figure out which other points are reachable by  
  a. Traveling a distance of 2x (as the crow flies), OR  
  b. Traveling less than x miles to an airport, taking one flight, and traveling less than x miles to another point.  
    Use x = 75 miles and x = 100 miles.  
```{r}
# Sequence of actions
#   (1) Find all grid points within `2x` distance of grid point
#   (2) Find all airports within `x` distance of grid point
#   (3) Find all connecting airports
#         -> `direct_from` object
#   (4) Find all points within connecting airport
#   (5) Calculate proportion of unique grid points reachable

# (1) FUNCTION : All grid points within `2x` distance of grid point
compute_points_within = function(x){
  points_within = vector("list", length = n_points)
  for(point in 1:n_points){
    x1 = lonlat_usa[point,]
    x2 = lonlat_usa
    dists = rdist.earth(x1, x2, miles = TRUE)
    within = which(dists <= x)
    points_within[[point]] = within
  }
  return(points_within)
}

# (2) FUNCTION : All airports within `x` distance of grid point
compute_airports_within = function(x){
  airports_within = vector("list", length = n_points)
  for(point in 1:n_points){
    x1 = lonlat_usa[point,]
    x2 = airport_data[, c(2, 3)]
    dists = rdist.earth(x1, x2, miles = TRUE)
    within = which(dists <= x)
    airports_within[[point]] = airport_data[within, 'codes']
  }
  return(airports_within)
}

# (4) FUNCTION : All points reachable within `x` distance of airports
compute_points_within_airport = function(x){
  points_within_airport = vector("list", length = n_airports)
  names(points_within_airport) = airports
  for(airport in airports){
    airport_idx = which(airport_data$codes == airport)
    x1 = airport_data[airport_idx, c(2, 3)]
    x2 = lonlat_usa
    dists = rdist.earth(x1, x2, miles = TRUE)
    within = which(dists <= x)
    points_within_airport[[airport]] = within
  }
  return(points_within_airport)
}

# (1) All grid points within `2x` distance of grid point
points_within_150 = compute_points_within(150)
points_within_200 = compute_points_within(200)

# (2) All airports within `x` distance of grid point
airports_within_75 = compute_airports_within(75)
airports_within_100 = compute_airports_within(100)

# (3) All connecting airports
#       -> `direct_from` object; and calculated already

# (4) Reachable gridpoints `x` miles from airport
points_within_airport_75 = compute_points_within_airport(75)
points_within_airport_100 = compute_points_within_airport(100)

compute_reachable_points = function(point, points_within_2x, airports_within_x,
                              points_within_airport_x, direct_from){
  # Hard encoding; `n_points`
  #   -> Later compute from length of arguments
  proportions = numeric(n_points)
  # Reachbale gridpoints from gridpoint
  reachable_points = numeric()
  # (1) Points within `2x` miles of grid point
  points_within = points_within_2x[[point]]
  reachable_points = append(points_within, reachable_points)
  # (2) Airports within `x` miles of grid point
  airports_within = airports_within_x[[point]]
  for(airport in airports_within){
    # (3) Direct flights from airport
    direct_flights = direct_from[[airport]]
    for(flight in direct_flights){
      # (4) Reachable gridpoints `x` miles from airport
      points_within_airport = points_within_airport_x[[flight]]
      reachable_points = append(reachable_points, points_within_airport)
    }
  }
  # Unique gridpoints reachable
  reachable_points = unique(reachable_points)
  return(reachable_points)
}
```

2 (cont). Plot the points reachable from the gridpoint closest to Ithaca. Use x = 75 miles and x = 100 miles.  
What proportion of grid points are reachable from the Ithaca gridpoint?
```{r fig.height=4.5, fig.width=6}

# From google, ITH lonlat is: 76.5019° W, 42.4440° N
# get point in lonlat_usa is closest to ithaca
airports_nearby(data.frame(lon=-75.5019,lat = 42.444), 'ith', lonlat_usa, rownames(lonlat_usa),x = 11) # index 6402
ith_point = 6402

# compute reachable points from ithaca x = 75 
ith_reachable_points_75 = compute_reachable_points(ith_point, points_within_150, airports_within_75,points_within_airport_75, direct_from)

# map of points reachable from ithaca when x = 75
par(mar=c(0,0,0,0))
map('usa',col="#f2f2f2",fill=TRUE, bg="white", lwd=0.05,mar=rep(0,4),border=0, ylim=c(-80,80) )
points(x=lonlat_usa$lon[ith_reachable_points_75], y=lonlat_usa$lat[ith_reachable_points_75], col="slateblue", cex=0.25, pch=20)

# proportion of points reachable from ithaca when x = 75
ith_proportion_75 = length(ith_reachable_points_75) / n_points
paste('proportion of reachable gridpoints in the continental US from Ithaca is: ', ith_proportion_75, ' when x = 75 miles')

# compute reachable points from ithaca x = 100 
ith_reachable_points_100 = compute_reachable_points(ith_point, points_within_200, airports_within_100,
                                                   points_within_airport_100, direct_from)
# map of points reachable from ithaca when x = 100
par(mar=c(0,0,0,0))
map('usa',col="#f2f2f2",fill=TRUE, bg="white", lwd=0.05,mar=rep(0,4),border=0, ylim=c(-80,80) )
points(x=lonlat_usa$lon[ith_reachable_points_100], y=lonlat_usa$lat[ith_reachable_points_100], col="slateblue", cex=0.25, pch=20)

# proportion of points reachable from ithaca when x = 100
ith_proportion_100 = length(ith_reachable_points_100) / n_points
paste('proportion of reachable gridpoints in the continental US from Ithaca is: ', ith_proportion_100, ' when x = 100 miles')

```

3. Repeat the previous exercise for every point in lonlat_usa, for x = 75, x = 100 miles
```{r}
compute_proportion = function(points_within_2x, airports_within_x,
                              points_within_airport_x, direct_from){
  # Hard encoding; `n_points`
  #   -> Later compute from length of arguments
  proportions = numeric(n_points)
  for(point in 1:n_points){
    # Reachbale gridpoints from gridpoint
    reachable_points = numeric()
    # (1) Points within `2x` miles of grid point
    points_within = points_within_2x[[point]]
    reachable_points = append(points_within, reachable_points)
    # (2) Airports within `x` miles of grid point
    airports_within = airports_within_x[[point]]
    for(airport in airports_within){
      # (3) Direct flights from airport
      direct_flights = direct_from[[airport]]
      for(flight in direct_flights){
        # (4) Reachable gridpoints `x` miles from airport
        points_within_airport = points_within_airport_x[[flight]]
        reachable_points = append(reachable_points, points_within_airport)
      }
    }
    # Unique gridpoints reachable
    reachable_points = unique(reachable_points)
    # (5) Proportion of gridpoints reachable from gridpoint
    proportion = length(reachable_points) / n_points
    proportions[point] = proportion
  }
  return(proportions)
}

proportion_75 = data.frame(proportion = compute_proportion(points_within_150, airports_within_75,
                                           points_within_airport_75, direct_from))

proportion_100 = data.frame(proportion = compute_proportion(points_within_200, airports_within_100,
                                                            points_within_airport_100, direct_from))
```

3 (cont). Make plots of the proportion of reachable grid points from every point in lonlat_usa for x = 75 miles and x = 100 miles. What is the average proportion of reachable grid points for x = 75 and x = 100?

```{r fig.height=4, fig.width=4}
# Make plots of the proportion of reachable grid points from every point in lonlat_usa for x = 75 miles and x = 100 miles. 

ggplot(proportion_75, aes(x=proportion))+
  geom_histogram() +
  theme_bw() +
  xlab('Proportion of US gridpoints that can be reached') +
  ylab('Frequency') +
  ggtitle('Histogram of proportion of US gridpoints that can be reached, radius = 75 miles, n = 9063 points') +
  theme(text = element_text(size = 4))

ggplot(proportion_100, aes(x=proportion))+
  geom_histogram() +
  theme_bw() +
  xlab('Proportion of US gridpoints that can be reached') +
  ylab('Frequency') +
  ggtitle('Histogram of proportion of US gridpoints that can be reached, radius = 100 miles, n = 9063 points') + 
  theme(text = element_text(size = 4))
# Average proportion of reachable grid points for x = 75 and x = 100
prop_75_avg = mean(proportion_75$proportion)
paste('Mean proportion of reachable gridpoints in the continental US: ', prop_75_avg, ' when x = 75 miles')

prop_100_avg = mean(proportion_100$proportion)
paste('Mean proportion of reachable gridpoints in the continental US: ', prop_100_avg, ' when x = 100 miles')

```

4. Redesign networks separately under the following 3 scenarios:
a. The total number of direct connections must remain constant. In other words, if you add a connection, you must remove a different connection.
b. The total number of connections to and from each airport must remain constant.
c. You can add or subtract up to two connections from each airport, leaving the other connections unchanged, and keeping the total number of connections constant.  

Construction of these network objects is in inst/scripts/new_design.R and new_design_aux.R  
  
  For each redesigned network, make a map of the reachable grid points from Ithaca
```{r fig.height=4.5, fig.width=6}
ith_point = 6402
ith_reachable_points_a_75 = compute_reachable_points(ith_point, points_within_150, airports_within_75,points_within_airport_75, redesign_A75[[1]])
ith_reachable_points_a_100 = compute_reachable_points(ith_point, points_within_200, airports_within_100,points_within_airport_100, redesign_A100[[1]])
ith_reachable_points_b_75 = compute_reachable_points(ith_point, points_within_150, airports_within_75, points_within_airport_75, redesign_B75[[1]])
ith_reachable_points_b_100 = compute_reachable_points(ith_point, points_within_200, airports_within_100,points_within_airport_100, redesign_B100[[1]])
ith_reachable_points_c_75 = compute_reachable_points(ith_point, points_within_150, airports_within_75,points_within_airport_75, redesign_C75[[1]])
ith_reachable_points_c_100 = compute_reachable_points(ith_point, points_within_200, airports_within_100,points_within_airport_100, redesign_C100[[1]])

# map of points reachable from ithaca
par(mar=c(0,0,0,0))
map('usa',col="#f2f2f2",fill=TRUE, bg="white", lwd=0.05,mar=rep(0,4),border=0, ylim=c(-80,80) )
points(x=lonlat_usa$lon[ith_reachable_points_a_75], y=lonlat_usa$lat[ith_reachable_points_a_75], col="slateblue", cex=0.25, pch=20)

par(mar=c(0,0,0,0))
map('usa',col="#f2f2f2",fill=TRUE, bg="white", lwd=0.05,mar=rep(0,4),border=0, ylim=c(-80,80) )
points(x=lonlat_usa$lon[ith_reachable_points_a_100], y=lonlat_usa$lat[ith_reachable_points_a_100], col="slateblue", cex=0.25, pch=20)

par(mar=c(0,0,0,0))
map('usa',col="#f2f2f2",fill=TRUE, bg="white", lwd=0.05,mar=rep(0,4),border=0, ylim=c(-80,80) )
points(x=lonlat_usa$lon[ith_reachable_points_b_75], y=lonlat_usa$lat[ith_reachable_points_b_75], col="slateblue", cex=0.25, pch=20)

par(mar=c(0,0,0,0))
map('usa',col="#f2f2f2",fill=TRUE, bg="white", lwd=0.05,mar=rep(0,4),border=0, ylim=c(-80,80) )
points(x=lonlat_usa$lon[ith_reachable_points_b_100], y=lonlat_usa$lat[ith_reachable_points_b_100], col="slateblue", cex=0.25, pch=20)

par(mar=c(0,0,0,0))
map('usa',col="#f2f2f2",fill=TRUE, bg="white", lwd=0.05,mar=rep(0,4),border=0, ylim=c(-80,80) )
points(x=lonlat_usa$lon[ith_reachable_points_c_75], y=lonlat_usa$lat[ith_reachable_points_c_75], col="slateblue", cex=0.25, pch=20)

par(mar=c(0,0,0,0))
map('usa',col="#f2f2f2",fill=TRUE, bg="white", lwd=0.05,mar=rep(0,4),border=0, ylim=c(-80,80) )
points(x=lonlat_usa$lon[ith_reachable_points_c_100], y=lonlat_usa$lat[ith_reachable_points_c_100], col="slateblue", cex=0.25, pch=20)


return_ith_proportion = function(letter, r, reachable_points){
  proportion = length(reachable_points)/n_points
  return(cat('Network', letter, 'r =',r, 'miles:', '\n', 'The proportion of continental US gridpoints reachable from Ithaca is:', proportion))
}

return_ith_proportion('A', 75, ith_reachable_points_a_75)
return_ith_proportion('A', 100, ith_reachable_points_a_100)
return_ith_proportion('B', 75, ith_reachable_points_b_75)
return_ith_proportion('B', 100, ith_reachable_points_b_100)
return_ith_proportion('C', 75, ith_reachable_points_c_75)
return_ith_proportion('C', 100, ith_reachable_points_c_100)

```

4 (cont). For each redesigned network, make a map of the proportion of reachable grid points from every grid point
```{r fig.height=4.5, fig.width=8}
# for each redesigned network (A,B,C) for 75, 100 miles
# create list with lonlat_usa df as object [[1]], and proportions object from redesign as [[2]]
# map heatmap of proportions on USA

return_lonlat_prop_list = function(redesign_list){
  lonlat_prop = vector('list', length = 2)
  lonlat_prop[[1]] = lonlat_usa
  lonlat_prop[[2]] = redesign_list[[2]]
  names(lonlat_prop) = c('lonlat', 'prop')
  return(lonlat_prop)
}

# run func for each redesigned network
redesign_A75_lonlat_proportions = return_lonlat_prop_list(redesign_A75)
redesign_A100_lonlat_proportions = return_lonlat_prop_list(redesign_A100)
redesign_B75_lonlat_proportions = return_lonlat_prop_list(redesign_B75)
redesign_B100_lonlat_proportions = return_lonlat_prop_list(redesign_B100)
redesign_C75_lonlat_proportions = return_lonlat_prop_list(redesign_C75)
redesign_C100_lonlat_proportions = return_lonlat_prop_list(redesign_C100)

# plot each redesigned network
quilt.plot(redesign_A75_lonlat_proportions$lonlat, redesign_A75_lonlat_proportions$prop,
           main = 'Proportion of Continental US Reachable by travelling 75 miles, Network A', cex.main = 0.75)
US( add=TRUE, col="magenta", lwd=2)

quilt.plot(redesign_A100_lonlat_proportions$lonlat, redesign_A100_lonlat_proportions$prop,
           main = 'Proportion of Continental US Reachable by travelling 100 miles, Network A', cex.main = 0.75)
US( add=TRUE, col="magenta", lwd=2)

quilt.plot(redesign_B75_lonlat_proportions$lonlat, redesign_B75_lonlat_proportions$prop,
           main = 'Proportion of Continental US Reachable by travelling 75 miles, Network B', cex.main = 0.75)
US( add=TRUE, col="magenta", lwd=2)

quilt.plot(redesign_B100_lonlat_proportions$lonlat, redesign_B100_lonlat_proportions$prop,
           main = 'Proportion of Continental US Reachable by travelling 100 miles, Network B', cex.main = 0.75)
US( add=TRUE, col="magenta", lwd=2)

quilt.plot(redesign_C75_lonlat_proportions$lonlat, redesign_C75_lonlat_proportions$prop,
           main = 'Proportion of Continental US Reachable by travelling 75 miles, Network C', cex.main = 0.75)
US( add=TRUE, col="magenta", lwd=2)

quilt.plot(redesign_C100_lonlat_proportions$lonlat, redesign_C100_lonlat_proportions$prop,
           main = 'Proportion of Continental US Reachable by travelling 100 miles, Network C', cex.main = 0.75)
US( add=TRUE, col="magenta", lwd=2)
```

4 (cont). For each redesigned network, report the average number of reachable grid points
```{r}
report_mean_prop = function(letter, r, redesign_obj){
  return(cat('Network', letter, 'r =',r, 'miles:', '\n', 'The mean proportion of gridpoints reachable in the continental US is:', mean(redesign_obj[[2]])))
}

report_mean_prop('A', 75, redesign_A75)
report_mean_prop('A', 100, redesign_A100)
report_mean_prop('B', 75, redesign_B75)
report_mean_prop('B', 100, redesign_B100)
report_mean_prop('C', 75, redesign_C75)
report_mean_prop('C', 100,redesign_C100)
```

